\section{introduction}

Principal Component Analysis (PCA) is a method for finding the principal components of a set of data points. The principal components of a collection of points in the space $\mathbb{R}^n$ are a sequence of $n$ direction vectors, where the $i^{th}$ vector is the direction of a line that best fits the data while being orthogonal to the first $i-1$ vectors. A best fitting line is defined as one that minimizes the average squared distance from the points to the line. The principal components form an orthonormal basis of the space $\mathbb{R}^n$.

One can think of PCA as fitting a p-dimensional ellipsoid to the data, where each axis of the ellipsoid represents a principal component.

One of the main applications for PCA is in quantitative finance, where data of large quantities is being reduced to a minimum amount of principal components that still hold the most important information about the data. 

In specific we look at a modification of PCA which is the Robust Principal Component Analysis. Here our aim is to recover a low rank Matrix $L$ from highly corrupted measurements $M$.

We start with a data matrix $M \in \mathbb{R}^{n \times n}$ with corrupted entries and try to find a composition into a sum $M = L + S$, where $L$ is a matrix of low rank and $S$ is sparse, which means that a lot of entries are zero. $M$ can be for example a covariance matrix which is by nature symmetric and positive definite.

Due to slow computation the usual algorithms for PCA are inefficient in some applications e.g. finance, where instantaneous calculation might be required. Therefore the use of a neuronal network is seen to perform better and can once trained be used to reduce computation times for a given problem.

In this project our aim is to implement the algorithm "Denise", an algorithm to compute a robust PCA for semidefinite matrices via a deep neural network. We train Denise on a randomly generated dataset and evaluate the performance of the deep neural network on synthetic and real-world covariance matrices.

The advantage of a neural network compared to the standard methods is that once trained the neural network delivers a function that outputs the desired decomposition of a dataset instantaneously, where using standard methods like convex optimization are computationally expensive. The algorithms are matrix specific which means one has to use the algorithm for every new data which is therefore very slow especially when data becomes large. 

For example in finance one needs robust low rank estimation of covariance matrices of hundreds of assets instantaneously where the high performance of a neural network is very practical.

The results achieved by Denise are comparable to several state-of-the-art algorithms in terms of decomposition quality but outperforms all existing algorithms by computation time. This is as explained above achieved by resulting in a single evaluation function that takes a matrix as an input and outputs the desired composition.

We further compare Denise to other algorithms like Principal Component Pursuit (PCP) as seen in the following chapters.




